---
title: "Final Report for Red Wine Analysis"
author: "Ruiqiang Chen, Michael DeWitt, David Williams, Alex Vannoy"
date: "7/28/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
library(ggfortify)
load.project()

(filenames <- list.files("src", pattern="*.R", full.names=TRUE))

for( a in 1:length(filenames)){
  on.exit(filenames[a])
  source(paste0(filenames[a]))
}
```

## Introduction

The purpose of this document is to report the proposed statistical models for classification of red wine. 

## Description of Data
The data set provided is the Wine dataset from UC Irvine. It consists of `r nrow(red_wine_data) ` with a total of `r ncol(red_wine_data) ` predictors. These predictors include the following `r names(red_wine_data)` with the quality feature being associated with the judgement of the individual wine's quality. Quality is the feature of interest for the dataset as the vinter is interested in judging the wine's quality through objective means rather than todays subjective method of averaging the 1-10 point judgment of tastetesters. The distribution of these different criteria can be seen below:
```{r echo=FALSE, warning=FALSE}
red_wine_data_index <-red_wine_data%>% 
  mutate(id = seq.int(nrow(red_wine_data)))
as.data.frame(red_wine_data_index) %>% 
  melt(id.vars = "id") %>% 
  ggplot(aes(value))+facet_wrap(~variable, scales = "free_x")+geom_histogram()+
  geom_histogram(binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3)))
```


Reviewing the individual components there appears to be a slight irregularity with total free sulfur dioxide. This can be seen in the histogram of this variable.
```{r echo=FALSE}
as.data.frame(red_wine_data_factors) %>% 
  melt() %>%
  filter(variable == "free_sulfur_dioxide") %>% 
  ggplot(aes( value))+
  geom_histogram(bins = 30)+
  labs(
    title = "Histogram of Free Sulfur Dioxide",
    caption = "From UCI Wine Data Set"
  )+
  xlab("Free Sulfur Dioxide")
```
As well as the fit of thithat display high studentized residuals and leverage and thus should be considered for removal in the modeling process. These wines are 1080 and 1082.
```{r echo = false}
autoplot(fit2)+
  labs(title = "Total Sulfur Dioxide")
```


## Method

###Regression

In order to select the best fit regression model several different modeling methods were tested. These include LEast Squares Regression, Ridge Regression, Lasso Regression, Principle Components Regression and Partial Least Squares Regression. For each of these methods the quality integer was the value that the model was attempting to predict. The data was divided into two sets, a training set to train the model and a testing set for model validation. We will now go deeper in the model generation process for each of these different modeling types and methods.

####Least Squares
The least squares regression method that was tested was the best subset selection. The methodology used to determine the best subset model was to first run cross validation on the entire model in order to determine the number of predictors to include in the model. Once this analysis indicated that any added predictor after 7 variables were selected did not increase the accuracy of the model greatly using this cross validated method.  USing this information, the data was divided into the training and testing data sets and a best subset selection method was done on the training data using a 7 predictor maximum.  

#####Residual Analysis
Here we need to make some plots against of the fit vs predictors and fit vs prediction to cross off that we considered our residuals

####Ridge Regression

####Lasso Regression

####Principle Components Regression
Why did we choose the # we did PC
####Partial Least Squares Regression

#### Model Selection
```{r warning= FALSE, echo = FALSE}
print(MSE_plot)
```


#### Residual Analysis

###Classification

#### Model Selection

#### Residual Analysis

###Comparison of Models

##Discussion

##Conclusion
##Issues

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
